---
title: "Workflow"
author: "Volker Schmid"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow for 3D SIM nucleome analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bioimagetools)
library(nucim)
```

## Workflow example
# Installation

Install *nucim* using *install.packages()*. The package relies on the *EBImage* package, which is available on [bioconductor](https://www.bioconductor.org), hence you need to set the repositories accordingly:

```{r install, eval=FALSE, collapse=TRUE}
setRepositories(ind=c(1,2))
install.packages("nucim")
```

Start with loading the necessary libraries:
```{r libraries, collapse=TRUE}
library(bioimagetools)
library(nucim)
```

# Read image and mask from DAPI
In the following, we use an example dataset available online. Use the *readTIF* function in *bioimagetools* to load the RGB image.

```{r read, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
img = readTIF("http://ex.volkerschmid.de/cell.tif")
sections = dim(img)[4]
x = y = 0.0395
z = 0.125
blue = img[,,3,] 
```

The *dapimask* function automatically finds the kernel using the DAPI (blue) channel. This takes about 30 seconds (all computation times are using MacBookPro 2019, 2,4 GhZ Quad-Core i5, 16 GB RAM)

```{r read, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
mask = dapimask(blue, c(x,y,z)*dim(img)[1:3], thresh="auto")
```
# Chromatin compaction classification

Next we find compaction classes from the DAPI channel (Computation time around 180 sec.).

```{r classes, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
classes = classify(blue, mask, 7, beta=0.1, z=x/z)

tab<-table.n(classes, 7, percentage=TRUE)
barplot(tab, ylab="percentage", xlab="chromatin compaction level",col=heatmap7())
par(pty="s")
img(classes, z=16, col=heatmap7(), mask=mask)
writeTIF(classes, "classes.tif")
```



# Distances of chromatin compaction classes
```{r distances, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE, eval=FALSE}
classes<-readClassTIF("classes.tif")
system.time({
distances = nearestClassDistances(classes,  voxelsize=c(x,y,z), classes=7, cores=4L)
})
save(distances,file="distances.Rdata")
plotNearestClassDistances(distances, method="min",ylim=c(0,.1),qu=.01)
plotNearestClassDistances(distances, method="quantile",ylim=c(0,.22),qu=.01)
plotNearestClassDistances(distances, method="boxplot",ylim=c(0,1.5))
```

# Map spots to classes
```{r cic, cache=TRUE, collapse=TRUE, warning=FALSE}
red = img[,,1,]
green = img[,,2,]
cc1<-colors.in.classes(classes,red,green,mask,7,type="thresh",plot=TRUE,
                       col1="red",col2="green",thresh1=0.05,thresh2=0.05,
                       test="Wilcoxon",ylim=c(0,.5),
                       xlab="chromatin compaction levels",ylab="percentage")
cc2<-colors.in.classes(classes,red,green,mask,7,type="intensity",plot=TRUE,
                       col1="red",col2="green",thresh1=0.05,thresh2=0.05,
                       test="Wilcoxon",ylim=c(0,.5),
                       xlab="chromatin compaction levels",ylab="percentage")
```
```{r spots, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
spots<-spots.combined(red=red,green=green,mask=mask,size=c(x,y,z),
                      full.voxel=FALSE,thresh.offset=0.05)
```
```{r cicspots, cache=TRUE, collapse=TRUE, warning=FALSE}
cc<-colors.in.classes(classes,spots$red,spots$green,mask,7,type="i",plot=TRUE,
                      col1="red",col2="green",test="Wilcoxon",ylim=c(0,.8),
                      xlab="chromatin compaction levels",ylab="percentage")
```
