---
title: "Workflow"
author: "Volker J Schmid"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow for 3D SIM nucleome analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bioimagetools)
library(nucim)
```

## Workflow example
# Installation

Install *nucim* using *install.packages()*. The package relies on the *EBImage* package, which is available on [bioconductor](https://www.bioconductor.org), hence you need to set the repositories accordingly:

```{r install, eval=FALSE, collapse=TRUE}
setRepositories(ind=c(1,2))
install.packages("nucim")
```

Start with loading the necessary libraries:
```{r libraries, collapse=TRUE}
library(bioimagetools)
library(nucim)
```

# Read image and mask from DAPI
In the following, we use an example dataset available online. Use the *readTIF* function in *bioimagetools* to load the RGB image.

```{r read, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
img = readTIF("http://ex.volkerschmid.de/cell.tif")
sections = dim(img)[4]
x = y = 0.0395
z = 0.125
blue = img[,,3,] 
```

The *dapimask* function automatically finds the kernel using the DAPI (blue) channel. This takes about 30 seconds (all computation times are using MacBookPro 2019, 2,4 GhZ Quad-Core i5, 16 GB RAM)

```{r mask, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
mask = dapimask(blue, c(x,y,z)*dim(img)[1:3], thresh="auto")
```
# Chromatin compaction classification

Next we find compaction classes from the DAPI channel (Computation time around 180 sec.).

```{r classes, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
classes = classify(blue, mask, 7, beta=0.1, z=x/z)
```

Barplot of classes and heatmap, finally save the classes image.

```{r plotclasses, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
tab<-table.n(classes, 7, percentage=TRUE)
barplot(tab, ylab="percentage", xlab="chromatin compaction level",col=heatmap7())
par(pty="s")
img(classes, z=16, col=heatmap7(), mask=mask)
writeTIF(classes, "classes.tif")
```


# Distances of chromatin compaction classes

Find the distances between compaction classes (computation time around 480 seconds, due to parallelization this depends on number of cores). The plotting function defines what we mean by distance: the minimum distance between class voxels, a quantile or plot a boxplot.

```{r distances, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE, eval=FALSE}
classes<-readClassTIF("classes.tif")
distances = nearestClassDistances(classes,  voxelsize=c(x,y,z), classes=7, cores=4L)
save(distances,file="distances.Rdata")
plotNearestClassDistances(distances, method="min",ylim=c(0,.1),qu=.01)
plotNearestClassDistances(distances, method="quantile",ylim=c(0,.22),qu=.01)
plotNearestClassDistances(distances, method="boxplot",ylim=c(0,1.5))
```

# Map spots to classes

We can compute the distribution of other color channels on compaction classes. Here we use a threshold approach and an intensity-weighted approach for comparison. A test on independence of color channels and compaction classes is automatically done (computation  time around 4 sec.).

```{r cic, cache=TRUE, collapse=TRUE, warning=FALSE}
red = img[,,1,]
green = img[,,2,]
cc1<-colors.in.classes(classes,red,green,mask,7,type="thresh",plot=TRUE,
                       col1="red",col2="green",thresh1=0.05,thresh2=0.05,
                       test="Wilcoxon",ylim=c(0,.5),
                       xlab="chromatin compaction levels",ylab="percentage")
cc2<-colors.in.classes(classes,red,green,mask,7,type="intensity",plot=TRUE,
                       col1="red",col2="green",thresh1=0.05,thresh2=0.05,
                       test="Wilcoxon",ylim=c(0,.5),
                       xlab="chromatin compaction levels",ylab="percentage")
```

Alternatively, we can find spots first and then use this to find the distribution on compaction classes (computation time *spots()* around 90 sec.). 

```{r spots, cache=TRUE, echo=TRUE, warning=FALSE, collapse=TRUE}
system.time({
spots<-spots.combined(red=red,green=green,mask=mask,size=c(x,y,z),
                      full.voxel=FALSE,thresh.offset=0.05)
})
```
```{r cicspots, cache=TRUE, collapse=TRUE, warning=FALSE}
cc<-colors.in.classes(classes,spots$red,spots$green,mask,7,type="i",plot=TRUE,
                      col1="red",col2="green",test="Wilcoxon",ylim=c(0,.8),
                      xlab="chromatin compaction levels",ylab="percentage")
```
